---
title: "Trade Intensity and Liquidity"
author: "Matt Brigida and William Pratt"
date: 
output:
  revealjs::revealjs_presentation:
    theme: league
    highlight: pygments
    center: true
---


## Goal of the Analysis

<p align="left">This analysis investigates how <b class="fragment highlight-green">liquidity is affected by periods of high trade intensity</b>.  We test whether the following measures, in the few seconds before a trade, have an effect on the book's liquidity in the <b class="fragment highlight-green">100 milliseconds</b> after the trade.</p>

-  <b class="fragment highlight-current-blue">number of changes in the orderbook.</b>

-  <b class="fragment highlight-current-blue">size of the bid-ask spread.</b>

-  <b class="fragment highlight-current-blue">number of trades.</b>

## Results

<p align="left">We find evidence consistent with larger bid-ask spreads leading to greater amounts of liquidity being provided by HFT post-trade and HFT providing liquidity when there is more activity in the orderbook.</p>

-  <b class="fragment highlight-current-blue">the activity in the orderbook prior to the trade.</b>


## Bid Side After a Sell Trade

```{r echo=FALSE}
knitr::include_app("https://mattbrigida.shinyapps.io/sell_bid/", height = "500px")
```


## Offer Side After a Sell Trade

```{r echo=FALSE}
knitr::include_app("https://mattbrigida.shinyapps.io/sell_offer/", height = "500px")
```


## Offer Side After a Buy Trade

```{r echo=FALSE}
knitr::include_app("https://mattbrigida.shinyapps.io/buy_offer/", height = "500px")
```


## Bid Side After a Buy Trade

```{r echo=FALSE}
knitr::include_app("https://mattbrigida.shinyapps.io/buy_bid/", height = "500px")
```

<!-- <section> -->
<!-- <div class="stretch"> -->
<!-- <iframe width="560" height="315" src="https://mattbrigida.shinyapps.io/buy_bid/" frameborder="0" allowfullscreen></iframe> -->
<!-- </div> -->
<!-- </section> -->

## Offer Side After a NA Trade

```{r echo=FALSE}
knitr::include_app("https://mattbrigida.shinyapps.io/na_offer/", height = "500px")
```

## Bid Side After a NA Trade

```{r echo=FALSE}
knitr::include_app("https://mattbrigida.shinyapps.io/na_bid/", height = "500px")
```



## 

<font size="5">

```{r echo = FALSE, message = FALSE, results = 'asis'}
load("../sept16multiplecontracts.RData")

names(buy.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(buy.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")

library(stargazer)

stargazer(buy.offer.lm, buy.bid.lm, sell.offer.lm, sell.bid.lm, na.offer.lm, na.bid.lm, covariate.labels = c("Num. Ch.", "Avg. BO", "Num. T"), dep.var.labels = c("Buy/Bid", "Buy/Offer", "Sell/Bid", "Sell/Offer", "NA/Bid", "NA/Offer"), digits = 2, no.space=FALSE, header=FALSE, type='html', omit.stat=c("LL","ser","f"), title = "Regression results for Sept. 16, 2013. The results shown are estimated using an interval of 1 second before the trade, and 100ms after the trade.", font.size="scriptsize", single.row = TRUE, column.sep.width = "1pt")

```
</font>

##
<font size="5">

```{r echo = FALSE, message = FALSE, results = 'asis'}
load("../sept17multiplecontracts.RData")

names(buy.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(buy.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")

stargazer(buy.offer.lm, buy.bid.lm, sell.offer.lm, sell.bid.lm, na.offer.lm, na.bid.lm, covariate.labels = c("Num. Ch.", "Avg. BO", "Num. T"), dep.var.labels = c("Buy/Bid", "Buy/Offer", "Sell/Bid", "Sell/Offer", "NA/Bid", "NA/Offer"), digits = 2, no.space=TRUE, header=FALSE, type='html', omit.stat=c("LL","ser","f"), title = "Regression results for Sept. 17, 2013. The results shown are estimated using an interval of 1 second before the trade, and 100ms after the trade.")

```
</font>

##
<font size="5">

```{r echo = FALSE, message = FALSE, results = 'asis'}
load("../sept18multiplecontracts.RData")

names(buy.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(buy.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")

stargazer(buy.offer.lm, buy.bid.lm, sell.offer.lm, sell.bid.lm, na.offer.lm, na.bid.lm, covariate.labels = c("Num. Ch.", "Avg. BO", "Num. T"), dep.var.labels = c("Buy/Bid", "Buy/Offer", "Sell/Bid", "Sell/Offer", "NA/Bid", "NA/Offer"), digits = 2, no.space=TRUE, header=FALSE, type='html', omit.stat=c("LL","ser","f"), title = "Regression results for Sept. 18, 2013. The results shown are estimated using an interval of 1 second before the trade, and 100ms after the trade.")

```
</font>

##
<font size="5">

```{r echo = FALSE, message = FALSE, results = 'asis'}
load("../sept19multiplecontracts.RData")

names(buy.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(buy.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")

stargazer(buy.offer.lm, buy.bid.lm, sell.offer.lm, sell.bid.lm, na.offer.lm, na.bid.lm, covariate.labels = c("Num. Ch.", "Avg. BO", "Num. T"), dep.var.labels = c("Buy/Bid", "Buy/Offer", "Sell/Bid", "Sell/Offer", "NA/Bid", "NA/Offer"), digits = 2, no.space=TRUE, header=FALSE, type='html', omit.stat=c("LL","ser","f"), title = "Regression results for Sept. 19, 2013. The results shown are estimated using an interval of 1 second before the trade, and 100ms after the trade.")

```
</font>

##
<font size="5">

```{r echo = FALSE, message = FALSE, results = 'asis'}
load("../sept20multiplecontracts.RData")

names(buy.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(buy.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")

stargazer(buy.offer.lm, buy.bid.lm, sell.offer.lm, sell.bid.lm, na.offer.lm, na.bid.lm, covariate.labels = c("Num. Ch.", "Avg. BO", "Num. T"), dep.var.labels = c("Buy/Bid", "Buy/Offer", "Sell/Bid", "Sell/Offer", "NA/Bid", "NA/Offer"), digits = 2, no.space=TRUE, header=FALSE, type='html', omit.stat=c("LL","ser","f"), title = "Regression results for Sept. 20, 2013. The results shown are estimated using an interval of 1 second before the trade, and 100ms after the trade.")

```
</font>

##
<font size="5">

```{r echo = FALSE, message = FALSE, results = 'asis'}
load("../sept23multiplecontracts.RData")

names(buy.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(buy.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")

stargazer(buy.offer.lm, buy.bid.lm, sell.offer.lm, sell.bid.lm, na.offer.lm, na.bid.lm, covariate.labels = c("Num. Ch.", "Avg. BO", "Num. T"), dep.var.labels = c("Buy/Bid", "Buy/Offer", "Sell/Bid", "Sell/Offer", "NA/Bid", "NA/Offer"), digits = 2, no.space=TRUE, header=FALSE, type='html', omit.stat=c("LL","ser","f"), title = "Regression results for Sept. 23, 2013. The results shown are estimated using an interval of 1 second before the trade, and 100ms after the trade.")

```
</font>

##
<font size="5">

```{r echo = FALSE, message = FALSE, results = 'asis'}
load("../sept24multiplecontracts.RData")

names(buy.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(buy.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")

stargazer(buy.offer.lm, buy.bid.lm, sell.offer.lm, sell.bid.lm, na.offer.lm, na.bid.lm, covariate.labels = c("Num. Ch.", "Avg. BO", "Num. T"), dep.var.labels = c("Buy/Bid", "Buy/Offer", "Sell/Bid", "Sell/Offer", "NA/Bid", "NA/Offer"), digits = 2, no.space=TRUE, header=FALSE, type='html', omit.stat=c("LL","ser","f"), title = "Regression results for Sept. 24, 2013. The results shown are estimated using an interval of 1 second before the trade, and 100ms after the trade.")

```
</font>

##
<font size="5">

```{r echo = FALSE, message = FALSE, results = 'asis'}
load("../sept25multiplecontracts.RData")

names(buy.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(buy.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")

stargazer(buy.offer.lm, buy.bid.lm, sell.offer.lm, sell.bid.lm, na.offer.lm, na.bid.lm, covariate.labels = c("Num. Ch.", "Avg. BO", "Num. T"), dep.var.labels = c("Buy/Bid", "Buy/Offer", "Sell/Bid", "Sell/Offer", "NA/Bid", "NA/Offer"), digits = 2, no.space=TRUE, header=FALSE, type='html', omit.stat=c("LL","ser","f"), title = "Regression results for Sept. 25, 2013. The results shown are estimated using an interval of 1 second before the trade, and 100ms after the trade.")

```
</font>

##
<font size="5">

```{r echo = FALSE, message = FALSE, results = 'asis'}
load("../sept26multiplecontracts.RData")

names(buy.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(buy.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")

stargazer(buy.offer.lm, buy.bid.lm, sell.offer.lm, sell.bid.lm, na.offer.lm, na.bid.lm, covariate.labels = c("Num. Ch.", "Avg. BO", "Num. T"), dep.var.labels = c("Buy/Bid", "Buy/Offer", "Sell/Bid", "Sell/Offer", "NA/Bid", "NA/Offer"), digits = 2, no.space=TRUE, header=FALSE, type='html', omit.stat=c("LL","ser","f"), title = "Regression results for Sept. 26, 2013. The results shown are estimated using an interval of 1 second before the trade, and 100ms after the trade.")
```
</font>

##
<font size="5">

```{r echo = FALSE, message = FALSE, results = 'asis'}
load("../sept27multiplecontracts.RData")

names(buy.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(buy.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")

stargazer(buy.offer.lm, buy.bid.lm, sell.offer.lm, sell.bid.lm, na.offer.lm, na.bid.lm, covariate.labels = c("Num. Ch.", "Avg. BO", "Num. T"), dep.var.labels = c("Buy/Bid", "Buy/Offer", "Sell/Bid", "Sell/Offer", "NA/Bid", "NA/Offer"), digits = 2, no.space=TRUE, header=FALSE, type='html', omit.stat=c("LL","ser","f"), title = "Regression results for Sept. 27, 2013. The results shown are estimated using an interval of 1 second before the trade, and 100ms after the trade.")
```

</font>
