---
title: "Trade Intensity and Liquidity"
author: "Matt Brigida and William Pratt"
date: 
bibliography: references.bib
output:
  revealjs::revealjs_presentation:
    theme: league
    highlight: pygments
    center: true
---


## Goal of the Analysis

<p align="left">This analysis investigates how <b class="fragment highlight-green">liquidity is affected by periods of high trade intensity</b>.  We test whether the following measures, in the few seconds before a trade, have an effect on the book's liquidity in the <b class="fragment highlight-green">100 milliseconds</b> after the trade.</p>

-  <b class="fragment highlight-current-blue">number of changes in the orderbook.</b>

-  <b class="fragment highlight-current-blue">size of the bid-ask spread.</b>

-  <b class="fragment highlight-current-blue">number of trades.</b>

## Related Literature

-  @Hend13 report algorithmic traders consume liquidity when spreads narrow, but supply when spread widens.

- @Chor00 observe a relationship and influence of spread on liquidity and market depth is time-varying, noting substantial variability across trading days and over time.

-  @Biai95 finds liquidity offering increases when bid-ask spread widens and observes intraday patterns.

## Identifying HFT

<p align="left">High-frequency traders (HFT) are identified in our analysis by the speed of their reaction to the trade.

- Specifically, we defing HFT as those who can react within 1 to 100 milliseconds after an event.

-  The event we use is a trade. 

-  Our method is a modification of the method used in @Hasb13 who used an order cancellation as the event.

## Data

<p align="left">Data are Market Depth Data [[link]](http://www.cmegroup.com/confluence/display/EPICSANDBOX/Market%2BDepth) for natural gas futures (ticker: NG) purchased directly from the Chicago Mercantile Exchange (CME).</p>

-  Market Depth Data contains all market messages to and from the CME, and is time-stamped to the millisecond.

- Using this data we can recreate the NG orderbook up to 10 levels deep.

<p align="left">Our data set is for NGV3 from September 16, 2013 to September 27, 2016 inclusive.</p>

## Why NG?

<p align="left">We choose to use the NYMEX natural gas book in this analysis because:</p>

- Unlike stock, all trades and quotes for NG take place in this one central book.

-  Since all trading takes place on the same computer server, there is no delay in orders due to location.

-  NG is heavily traded and very volatile.

## Data Preparation

#. We first extract trades for a single contract with no trades during the same millisecond, and the following 100 ms.

## Method

<p align="left">We model the change in liquidity post-trade as functions of three pre-trade variables. Our model is:</p>

$\Delta L = \beta_0 + \beta_1 numC + \beta_2 avgBA + \beta_3 numT + \mu$

<p align="left">where:</p>

-  <span class="fragment highlight-current-blue">$avgBA$</span> is the average Bid-Ask spread

-  <span class="fragment highlight-current-blue">$numT$</span> is the number of trades

-  <span class="fragment highlight-current-blue">$numC$</span> is the number of changes in the orderbook

-  <span class="fragment highlight-current-blue">$y$</span> is the number of milliseconds prior to the trade.

## Method

<p align="left">We estimate this equation for each side of the orderbook, separately for buy and sell trades, and also for trades where the initiator is unknown.  We thus have 6 separate regressions.</p>

## Method

<p align="left">we calculate $avgBA$, $numT$, and $numC$ over 30 different pre-trade intervals ranging from 100 ms, to 3000 ms prior to the trade.  The end point of each interval was 1 ms prior to the trade.</p>

<p align="left">Similarly, we calculate $Delta L$ over intervals starting 1 ms after the trade, incrementing by 1 ms, until 100 ms after the trade.</p>

-  This gives us a total of $6(3000) = 18000$ individual regressions for each trading day.

## 3D Plots

<p align="left">Following are 3-dimensional plots of the coefficients estimated over the range of possible pre and post-trade intervals.</p>

-  We use these plots to investigate how sensitive our coefficient estimate is to our choice of pre and post-trade interval.  

-  We find an interval of 1 second before the trade, and 100ms after the trade, affords a stable estimate of the coefficient.

## Bid Side After a Sell Trade

```{r echo=FALSE}
knitr::include_app("https://mattbrigida.shinyapps.io/sell_bid/", height = "500px")
```


## Offer Side After a Sell Trade

```{r echo=FALSE}
knitr::include_app("https://mattbrigida.shinyapps.io/sell_offer/", height = "500px")
```


## Offer Side After a Buy Trade

```{r echo=FALSE}
knitr::include_app("https://mattbrigida.shinyapps.io/buy_offer/", height = "500px")
```


## Bid Side After a Buy Trade

```{r echo=FALSE}
knitr::include_app("https://mattbrigida.shinyapps.io/buy_bid/", height = "500px")
```

<!-- <section> -->
<!-- <div class="stretch"> -->
<!-- <iframe width="560" height="315" src="https://mattbrigida.shinyapps.io/buy_bid/" frameborder="0" allowfullscreen></iframe> -->
<!-- </div> -->
<!-- </section> -->

## Offer Side After a NA Trade

```{r echo=FALSE}
knitr::include_app("https://mattbrigida.shinyapps.io/na_offer/", height = "500px")
```

## Bid Side After a NA Trade

```{r echo=FALSE}
knitr::include_app("https://mattbrigida.shinyapps.io/na_bid/", height = "500px")
```

## 

<font size="5">

```{r echo = FALSE, message = FALSE, results = 'asis'}
load("../sept16multiplecontracts.RData")

names(buy.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(buy.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")

library(stargazer)

stargazer(buy.offer.lm, buy.bid.lm, sell.offer.lm, sell.bid.lm, na.offer.lm, na.bid.lm, covariate.labels = c("Num. Ch.", "Avg. BO", "Num. T"), dep.var.labels = c("Buy/Bid", "Buy/Offer", "Sell/Bid", "Sell/Offer", "NA/Bid", "NA/Offer"), digits = 2, no.space=FALSE, header=FALSE, type='html', omit.stat=c("LL","ser","f"), title = "Regression results for Sept. 16, 2013. The results shown are estimated using an interval of 1 second before the trade, and 100ms after the trade.", font.size="scriptsize", single.row = TRUE, column.sep.width = "1pt")

```
</font>

##
<font size="5">

```{r echo = FALSE, message = FALSE, results = 'asis'}
load("../sept17multiplecontracts.RData")

names(buy.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(buy.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")

stargazer(buy.offer.lm, buy.bid.lm, sell.offer.lm, sell.bid.lm, na.offer.lm, na.bid.lm, covariate.labels = c("Num. Ch.", "Avg. BO", "Num. T"), dep.var.labels = c("Buy/Bid", "Buy/Offer", "Sell/Bid", "Sell/Offer", "NA/Bid", "NA/Offer"), digits = 2, no.space=TRUE, header=FALSE, type='html', omit.stat=c("LL","ser","f"), title = "Regression results for Sept. 17, 2013. The results shown are estimated using an interval of 1 second before the trade, and 100ms after the trade.")

```
</font>

##
<font size="5">

```{r echo = FALSE, message = FALSE, results = 'asis'}
load("../sept18multiplecontracts.RData")

names(buy.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(buy.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")

stargazer(buy.offer.lm, buy.bid.lm, sell.offer.lm, sell.bid.lm, na.offer.lm, na.bid.lm, covariate.labels = c("Num. Ch.", "Avg. BO", "Num. T"), dep.var.labels = c("Buy/Bid", "Buy/Offer", "Sell/Bid", "Sell/Offer", "NA/Bid", "NA/Offer"), digits = 2, no.space=TRUE, header=FALSE, type='html', omit.stat=c("LL","ser","f"), title = "Regression results for Sept. 18, 2013. The results shown are estimated using an interval of 1 second before the trade, and 100ms after the trade.")

```
</font>

##
<font size="5">

```{r echo = FALSE, message = FALSE, results = 'asis'}
load("../sept19multiplecontracts.RData")

names(buy.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(buy.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")

stargazer(buy.offer.lm, buy.bid.lm, sell.offer.lm, sell.bid.lm, na.offer.lm, na.bid.lm, covariate.labels = c("Num. Ch.", "Avg. BO", "Num. T"), dep.var.labels = c("Buy/Bid", "Buy/Offer", "Sell/Bid", "Sell/Offer", "NA/Bid", "NA/Offer"), digits = 2, no.space=TRUE, header=FALSE, type='html', omit.stat=c("LL","ser","f"), title = "Regression results for Sept. 19, 2013. The results shown are estimated using an interval of 1 second before the trade, and 100ms after the trade.")

```
</font>

##
<font size="5">

```{r echo = FALSE, message = FALSE, results = 'asis'}
load("../sept20multiplecontracts.RData")

names(buy.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(buy.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")

stargazer(buy.offer.lm, buy.bid.lm, sell.offer.lm, sell.bid.lm, na.offer.lm, na.bid.lm, covariate.labels = c("Num. Ch.", "Avg. BO", "Num. T"), dep.var.labels = c("Buy/Bid", "Buy/Offer", "Sell/Bid", "Sell/Offer", "NA/Bid", "NA/Offer"), digits = 2, no.space=TRUE, header=FALSE, type='html', omit.stat=c("LL","ser","f"), title = "Regression results for Sept. 20, 2013. The results shown are estimated using an interval of 1 second before the trade, and 100ms after the trade.")

```
</font>

##
<font size="5">

```{r echo = FALSE, message = FALSE, results = 'asis'}
load("../sept23multiplecontracts.RData")

names(buy.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(buy.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")

stargazer(buy.offer.lm, buy.bid.lm, sell.offer.lm, sell.bid.lm, na.offer.lm, na.bid.lm, covariate.labels = c("Num. Ch.", "Avg. BO", "Num. T"), dep.var.labels = c("Buy/Bid", "Buy/Offer", "Sell/Bid", "Sell/Offer", "NA/Bid", "NA/Offer"), digits = 2, no.space=TRUE, header=FALSE, type='html', omit.stat=c("LL","ser","f"), title = "Regression results for Sept. 23, 2013. The results shown are estimated using an interval of 1 second before the trade, and 100ms after the trade.")

```
</font>

##
<font size="5">

```{r echo = FALSE, message = FALSE, results = 'asis'}
load("../sept24multiplecontracts.RData")

names(buy.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(buy.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")

stargazer(buy.offer.lm, buy.bid.lm, sell.offer.lm, sell.bid.lm, na.offer.lm, na.bid.lm, covariate.labels = c("Num. Ch.", "Avg. BO", "Num. T"), dep.var.labels = c("Buy/Bid", "Buy/Offer", "Sell/Bid", "Sell/Offer", "NA/Bid", "NA/Offer"), digits = 2, no.space=TRUE, header=FALSE, type='html', omit.stat=c("LL","ser","f"), title = "Regression results for Sept. 24, 2013. The results shown are estimated using an interval of 1 second before the trade, and 100ms after the trade.")

```
</font>

##
<font size="5">

```{r echo = FALSE, message = FALSE, results = 'asis'}
load("../sept25multiplecontracts.RData")

names(buy.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(buy.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")

stargazer(buy.offer.lm, buy.bid.lm, sell.offer.lm, sell.bid.lm, na.offer.lm, na.bid.lm, covariate.labels = c("Num. Ch.", "Avg. BO", "Num. T"), dep.var.labels = c("Buy/Bid", "Buy/Offer", "Sell/Bid", "Sell/Offer", "NA/Bid", "NA/Offer"), digits = 2, no.space=TRUE, header=FALSE, type='html', omit.stat=c("LL","ser","f"), title = "Regression results for Sept. 25, 2013. The results shown are estimated using an interval of 1 second before the trade, and 100ms after the trade.")

```
</font>

##
<font size="5">

```{r echo = FALSE, message = FALSE, results = 'asis'}
load("../sept26multiplecontracts.RData")

names(buy.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(buy.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")

stargazer(buy.offer.lm, buy.bid.lm, sell.offer.lm, sell.bid.lm, na.offer.lm, na.bid.lm, covariate.labels = c("Num. Ch.", "Avg. BO", "Num. T"), dep.var.labels = c("Buy/Bid", "Buy/Offer", "Sell/Bid", "Sell/Offer", "NA/Bid", "NA/Offer"), digits = 2, no.space=TRUE, header=FALSE, type='html', omit.stat=c("LL","ser","f"), title = "Regression results for Sept. 26, 2013. The results shown are estimated using an interval of 1 second before the trade, and 100ms after the trade.")
```
</font>

##
<font size="5">

```{r echo = FALSE, message = FALSE, results = 'asis'}
load("../sept27multiplecontracts.RData")

names(buy.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(buy.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(sell.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.bid.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")
names(na.offer.lm$coefficients) <- c("Intercept", "Number Changes", "Avg. Bid/Offer", "Number Trades")

stargazer(buy.offer.lm, buy.bid.lm, sell.offer.lm, sell.bid.lm, na.offer.lm, na.bid.lm, covariate.labels = c("Num. Ch.", "Avg. BO", "Num. T"), dep.var.labels = c("Buy/Bid", "Buy/Offer", "Sell/Bid", "Sell/Offer", "NA/Bid", "NA/Offer"), digits = 2, no.space=TRUE, header=FALSE, type='html', omit.stat=c("LL","ser","f"), title = "Regression results for Sept. 27, 2013. The results shown are estimated using an interval of 1 second before the trade, and 100ms after the trade.")
```

</font>

## Results

<p align="left">Some salient features of the results follow.  On the left are the pre-trade measures, and on the right is post-trade liquidity.</p>

-  <b class="fragment highlight-current-blue">More trades $\Rightarrow$ less liquidity (part. on opp. side).</b>

-  <b class="fragment highlight-current-blue">More book changes $\Rightarrow$ more liquidity (part. on opp. side).</b>

-  <b class="fragment highlight-current-blue">The above is true on *both* sides for NA trades.</b>

-  <b class="fragment highlight-current-blue">The evidence for the impact of the size of the bid-offer spread is *mixed*.</b>

## Trades and Liquidity

<p align="left">Trades take liquidity, and so our result may be interpreted as if there is more liquidity taken prior to the trade, the less liquidity is provided after.</p>

## Book Changes and the Spread

<p align="left">This is evidence that activity in the orderbook may be a more important determinant of liquidity provision than the size of the spread.</p> 
-  HFT is more willing to provide liquidity when the book is active, and not simply because the compensation for providing liquidity is high.

## Contact

matthew.brigida [at] sunyit.edu

## References
